Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex

In $2011$, the International Organization for Standardization (ISO) issued an
update for software product quality ISO/IEC 25010~\cite{iso:2011} considering
\emph{Security} as one of the main software product quality characteristics.
Unmaintainable code is hard to test, analyze and re-use, and thus more difficult
to find vulnerabilities and defects. Therefore, writing maintainable code
prevents the introduction of vulnerabilities in the codebases. In this work, we are
concerned about the impact that patching vulnerabilities may have in the
open-source software maintainability. Are developers affecting software
maintainability when trying to improve software safety? We suspect that some of
the patches may have a negative impact on the software maintainability and
possibly even be the cause of the introduction of new vulnerabilities.

ISO describes software maintainability as ``the degree of effectiveness and
efficiency with which a software product or system can be modified to improve
it, correct it or adapt it to changes in environment, and in requirements'' on
software quality ISO/IEC 25010. As ISO does not provide any specific guidelines or
formula to calculate maintainability, we resort on Software Improvement
Group (SIG)'s web-based source code analysis service Better Code Hub
(BCH)~\cite{Visser:2016:OREILLY} to compute the maintainability of a given
project. There are other well-known standards and models that have been proposed to increase software safety:
Common Criteria~\cite{common:2009} which suffered a lot of criticism regarding the
costs associated and poor technical evaluation; the OWASP
Application Security Verification Criteria~\cite{oswap:2009} which is focused only
on web applications, and a model proposed by Xu et al. ($2013$)~\cite{6616351} for
rating software security which we believe it was one of the first steps taken
by SIG to introduce security on their maintainability model.

@@

\end{lstlisting}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The methodology
comprehends the following steps, as illustrated in Figure~\ref{fig:met}:
\begin{enumerate}
	\item Extract relevant data from a dataset containing
$716$ security refactorings collected from open-source software available on
GitHub~\cite{Reis:2017:IJSSE}.
	\item A baseline of regular commits was
randomly collected from the list of projects of the main dataset to evaluate the impact of regular commits on the maintainability of open-source
software.
\item Use the Software Improvement Group's\footnote{SIG's website is available
at https://www.sig.eu/ (Accessed on \today{})} web-based source code analysis
service \emph{Better Code Hub} (BCH)\footnote{BCH's website is available at
https://bettercodehub.com/ (Accessed on \today{})} to quantify
maintainability for both security and regular commits.
\end{enumerate}

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}
%
We found that $40.2\%$ of the security refactoring have a negative impact on the
maintainability of open-source software, Figure~\ref{fig:secvsreg}. This raises
a new tradeoff when developers need to fix vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of safety. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure and user-friendly~\cite{5287006, 7676144}. Unfortunately,
	there are several vulnerabilities that stay undisclosed. Thus, developers should not
	trust them blindly and check defaults.
	\item \textbf{Documentation} of software should provide developers with the best practices to implement safe and maintainable software. Unmaintainable code is difficult to test, analyze and re-use.
	\item \textbf{Proper Authorization Mechanisms} should be designed and implemented. This is
	one of the most complex security features to implement. Some mechanisms might assist authorization maintenance in long run, such as the combination of \texttt{HTTP} filters with \texttt{Spring Security} annotations.
	\item \textbf{Deprecated code} should be avoided by developers. They should integrate tools such as BCH on their projects CI/CD pipelines to help them be more aware of the maintainability of their code and issues (e.g., dead and duplicated code).
	\item\textbf{Programming Languages} should provide new mechanisms to easily implement security patterns without endangering maintainability.

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to the validity of this study.
%
\subsection{Construct}
%

The maintainability formula was inferred based on the BCH's reports.
The high amount of different projects and backgrounds may require different maintainability
standards. However, BCH does have a representative benchmark of closed and open-source
software projects that are under the calculation of the thresholds used in the computation
of each maintainability guideline (Table~\ref{tab:guidelines})~\cite{Visser:2016:OREILLY, Baggen2012}.

@@

\Luis{A primeira threat que falas é que os resultados podem nao generalizar
para software fechado. Eu acho que deviamos ser mais diretos.}
\Sof{Melhor?}

@@

\Luis{eu nao acho que usar $10$ guidelines proposed by the Software Improvement
Group (SIG) seja algo que nos esteja a diferenciar. Acho que o a prinicpal
diferenca é que estamos a estudar maintainability issues noutro contexto.}\Sof{Melhor?}

@@

\Luis{ tambem fizemos analise por linguagem, nesta ultima citacao deviamos falar sobre isso tb.}\Sof{Acho que esta!}

@@

\Luis{secure security frameworks, nao percebi bem o que se quer dizer aí.}
\Sof{Melhor?}

>>>>

In $2011$, the International Organization for Standardization (ISO) issued an
update for software product quality ISO/IEC 25010~\cite{iso:2011} considering
\emph{Security} as one of the main software product quality characteristics.
Unmaintainable code is hard to test, analyze and re-use, and thus more difficult
to find vulnerabilities and defects. Therefore, writing maintainable code
prevents the introduction of vulnerabilities in the codebases. In this work, we are
concerned about the impact that patching vulnerabilities may have in the
open-source software maintainability. Are developers affecting software
maintainability when trying to improve software security? We suspect that some
patches may have a negative impact on the software maintainability and
possibly even be the cause of the introduction of new vulnerabilities.

ISO describes software maintainability as ``the degree of effectiveness and
efficiency with which a software product or system can be modified to improve
it, correct it or adapt it to changes in environment, and in requirements'' on
software quality ISO/IEC 25010. As ISO does not provide any specific guidelines or
formula to calculate maintainability, we resort on Software Improvement
Group (SIG)'s web-based source code analysis service Better Code Hub
(BCH)~\cite{Visser:2016:OREILLY} to compute the maintainability of a given
project. There are other well-known standards and models that have been proposed to increase software security:
Common Criteria~\cite{common:2009} which suffered a lot of criticism regarding the
costs associated and poor technical evaluation; the OWASP
Application Security Verification Criteria~\cite{oswap:2009} which is focused only
on web applications, and a model proposed by Xu et al. ($2013$)~\cite{6616351} for
rating software security which we believe it was one of the first steps taken
by SIG to introduce security on their maintainability model.

@@

\end{lstlisting}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in this section. The
methodology comprehends the following steps, as illustrated in
Figure~\ref{fig:met}:

\begin{enumerate}
	\item Extract relevant data from a dataset containing $716$ security
	refactorings collected from open-source software available on
	GitHub~\cite{Reis:2017:IJSSE}.
	\item A baseline of regular commits was randomly collected from the list of
	projects of the main dataset to evaluate the impact of regular commits on the
	maintainability of open-source software.
  \item Use the Software Improvement Group's\footnote{SIG's website is available
  at https://www.sig.eu/ (Accessed on \today{})} web-based source code analysis
  service \emph{Better Code Hub} (BCH)\footnote{BCH's website is available at
  https://bettercodehub.com/ (Accessed on \today{})} to quantify maintainability
  for both security and regular commits.
\end{enumerate}

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}
%
We found that $40.2\%$ of the security refactoring have a negative impact on the
maintainability of open-source software, Figure~\ref{fig:secvsreg}. This raises
a new tradeoff when developers need to fix vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of security. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure and
	user-friendly~\cite{5287006, 7676144}. Unfortunately, there are several
	vulnerabilities that stay undisclosed. Thus, developers should not trust them
	blindly and check defaults.

	\item \textbf{Documentation} of software should provide developers with the
	best practices to implement secure and maintainable software. Unmaintainable
	code is difficult to test, analyze and re-use.

	\item \textbf{Proper Authorization Mechanisms} should be designed and
	implemented. This is one of the most complex security features to implement.
	Some mechanisms might assist authorization maintenance in long run, such as
	the combination of \texttt{HTTP} filters with \texttt{Spring Security}
	annotations.

	\item \textbf{Deprecated code} should be avoided by developers. They should
	integrate tools such as BCH on their projects CI/CD pipelines to help them be
	more aware of the maintainability of their code and issues (e.g., dead and
	duplicated code).

	\item\textbf{Programming Languages} should provide new mechanisms to easily
	implement security patterns without endangering maintainability.
\end{itemize}

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to the validity of this
study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the BCH's reports. The high
amount of different projects and backgrounds may require different
maintainability standards. However, BCH does have a representative benchmark of
closed and open-source software projects that are under the calculation of the
thresholds used in the computation of each maintainability guideline
(Table~\ref{tab:guidelines})~\cite{Visser:2016:OREILLY, Baggen2012}.

@@

@@

@@

@@
